NAME
       FFC - the FEniCS Form Compiler


SYNOPSIS
       ffc [-h] [-v] [-d debuglevel] [-s] [-l language] [-r representation]
       [-f option] [-O] [-o output-directory] [-q quadrature-rule] ...
       input.ufl ...


DESCRIPTION
       Compile multilinear forms into efficient low-level code.

       The FEniCS Form Compiler FFC accepts as input one or more  files,  each
       specifying  one or more multilinear forms, and compiles the given forms
       into efficent low-level code for automatic assembly of the tensors rep-
       resenting  the multilinear forms. In particular, FFC compiles a pair of
       bilinear and linear forms defining a variational problem into code that
       can be used to efficiently assemble the corresponding linear system.

       By  default, FFC generates code according to the UFC specification ver-
       sion 1.0 (Unified Form-assembly Code, see  http://www.fenics.org/)  but
       this  can  be  controlled  by  specifying  a  different output language
       (option -l). It is also possible to add new output languages to FFC.

       For a full description of FFC, including a specification  of  the  form
       language  used to define the multilinear forms, see the FFC user manual
       available on the FEniCS web page: http://www.fenics.org/


OPTIONS
       -h, --help
              Display help text and exit.

       -v, --version
              Display version number and exit.

       -d debuglevel, --debug debuglevel
              Specify debug level (default is 0).

       -s, --silent
              Silent mode, no output is printed (same as --debuglevel -1).

       -l language, --language language
              Specify output language, one of 'ufc' (default) or 'dolfin' (UFC
              with a small layer of DOLFIN-specific bindings).

       -r representation, --representation representation
              Specify  representation  for precomputation and code generation,
              one of 'tensor' (default) or 'quadrature' (experimental).

       -f option
              Specify code generation options. The list of options available
              depends on the specified  language  (format). Current  options  
              include  -fblas, -fno-foo, -fprecision=n,
              -fprecompute_basis_const, -fprecompute_ip_const,
              -fquadrature_degree=n and, -fsplit, described in detail below.

       -f blas
              Generate code that uses BLAS to compute tensor products.
              This option is currently ignored, but can be  used to reduce the
              code size when the BLAS option is (re-)implemented in future
              versions.

       -f no-foo
              Don't  generate code for UFC function with name 'foo'. Typical
              options include -fno-evaluate_basis and
              -fno-evaluate_basis_derivatives to reduce the size of the
              generated code when these functions are not needed.

       -f precision=n
              Set the number of significant digits to n in the generated code.
              The default value of n is 15.

       -f precompute_basis_const
              Additional optimisation option for quadrature representation.
              This option is ignored if optimisation is  not used 
              (see  -O  option),  and it also implies the precompute_ip_const
              option. This option will generate code that precompute terms
              which are constant in the loops involving basis indices. This can
              result in  a  reduction  of  the operation count and thereby
              improve the runtime efficiency of the generated code. However,
              the improvements depends on the GCC compiler options as well as
              the characteristics of the variational form.

       -f precompute_ip_const
              Like the precompute_basis_const option with the only difference
              that code will be generated to compute terms which are constant
              in the loops involving the integration points only.

       -f quadrature_degree=n
              Will  generate  a  quadrature  rule accurate up to degree n
              regardless of the polynomial degree of the form. This option is
              only valid for UFL forms and the specified degree will apply to
              ALL terms of the  given  form for  which no degree has been
              specified through metadata! As default FFC will determine the
              degree automatically from the form.

       -f split
              Generate separate files for declarations and the implementation.

       -O, --optimize
              Generate optimized code with a lower operation  count compared to
              non-optimized code for the assembly of  the local element tensor.
              This  will  in  general increase the  run-time performance of the
              code. If  the representation (see -r option) is 'tensor' then FFC
              will use FErari optimizations. However, this option is  currently
              ignored. This  option  requires  FErari  and  should be used with
              caution since it may be very costly (at compile-time)  for  other
              than  simple  forms.  If the representation is  'quadrature'  the
              compile-time increase tends to be  much  less drastic compared to
              FErari for very complex forms.

       -o directory, --output-directory directory
              Specify the directory where the generated files should be written
              to. The default output directory is the current ('.') directory.

       -q rule, --quadrature-rule rule
              Specify the quadrature rule that should be used when integrating
              the   forms.  This   will   affect  both  tensor and  quadrature
              representation.  Currently,  no  quadrature   rules   has   been
              implemented so the default from FIAT will be used.


       BUGS

       Send comments, questions, bug reports etc. to ffc-dev@fenics.org.

AUTHOR
       Written by Anders Logg (logg@simula.no) with help from Kristian {\O}lgaard,
       Marie Rognes, Garth N. Wells and many others.


                                                                        FFC(1)
